---
title: "Bars and restaurants"
author: "Andrew Ba Tran"
date: "9/10/2020"
output:
  html_document:
    code_folding: hide
---

This is a summary of all the cases per 10,000 people in states based on milestones such as reopening or closing or re-reopening.

Click on the "Code" buttons on the right to expand the underlying code used in calculations

```{r setup, warning=F, message=F}

if (!require("devtools")) {
  install.packages("devtools")
}

packages <- c("tidyverse", "knitr", "DT", "ggrepel", "geofacet")

if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())), repos = "http://cran.us.r-project.org")  
}


library(tidyverse)
library(knitr)
library(DT)
library(ggrepel)
library(geofacet)

# much of the data processing for this file can be found in 01_exploratory.R
for_export <- read_csv("data/for_export.csv")
```


# Correlation test

```{r correlation, warning=F, message=F}

states_bars <- for_export %>% 
  filter(bars_reopen_baseline %in% c(3, 4)) %>% 
  select(-detail, -detail_date) %>% 
  mutate(bar_per=bar_per*100) %>% 
  unique() %>% 
  select(region, NAME, MMWRyear, MMWRweek, cases, population) %>% 
  arrange(region, NAME, MMWRyear, MMWRweek) %>% 
  group_by(region, NAME, MMWRyear) %>% 
  mutate(id=row_number()) %>% 
  mutate(id=case_when(
    id == 1 ~ "week_a",
    id == 2 ~ "week_b"
  )) %>% 
  group_by(region, id) %>% 
  summarize(cases=sum(cases),
            population=sum(population)) %>% 
  mutate(cases_per_10k=cases/population*10000) %>% 
  select(-cases, -population) %>% 
  pivot_wider(names_from="id", values_from="cases_per_10k") %>% 
  mutate(cases_pc_increase=(week_b-week_a)/week_a*100)


states_bar_traffic <- for_export %>% 
  filter(bars_reopen_baseline %in% c(0, 1)) %>% 
  select(-detail, -detail_date) %>% 
  unique() %>% 
  select(region, NAME, MMWRyear, MMWRweek,bar_visits, total_devices) %>% 
  arrange(region, NAME, MMWRyear, MMWRweek) %>% 
  group_by(region, NAME, MMWRyear) %>% 
  mutate(id=row_number()) %>% 
  mutate(id=case_when(
    id == 1 ~ "bar_a",
    TRUE ~ "bar_b"
  )) %>% 
  group_by(region, id) %>% 
  summarize(bar_visits=sum(bar_visits),
            total_devices=mean(total_devices)) %>% 
  mutate(bar_traffic=bar_visits/total_devices*1000) %>% 
  select(-bar_visits, -total_devices) %>% 
  pivot_wider(names_from="id", values_from="bar_traffic") %>% 
  mutate(bar_traffic_increase=(bar_b-bar_a)/bar_a*100)

states_bars <- left_join(states_bars,  states_bar_traffic) 

ggplot(states_bars, aes(x=cases_pc_increase, y=bar_traffic_increase, label=region)) +
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal() +
  geom_text_repel() +
  labs(title="Bar foot traffic change after reopening vs. lagged change in covid cases",
       x="Change in cases per capita 3-4 weeks after bars reopened",
       y="Change in foot traffic 1 week after bars reopened")

cor.test(states_bars$cases_pc_increase, states_bars$bar_traffic_increase)

```

# Bars reopening as a baseline

```{r bars_baseline, warning=F, message=F, fig.width=9, fig.height=6.5}
four<- c(-4, -3, -2, -1, 0, 1,2,3, 4)

for_export %>% 
  #  filter(bars_baseline %in% three) %>% 
  filter(bars_reopen_baseline %in% four) %>% 
  select(-detail, -detail_date) %>% 
  mutate(bar_per=bar_per*100) %>% 
  unique() %>% 
  filter(bar_per<200) %>% 
  ggplot(aes(x=bars_reopen_baseline, y=cases_pc)) +
  geom_line() +
  geom_line(aes(x=bars_reopen_baseline, y=bar_per), color="blue") +
  geom_vline(aes(xintercept=0), alpha=.5, color="red")+
  facet_geo(~region) +
  labs(title="Cases per capita four weeks before, when bars reopened, and four weeks after")

```


# Visualizing when bars and restaurants closed or reopened

```{r exploratory,warning=F, message=F, fig.width=9, fig.height=6.5}


shutdown_baseline <- for_export %>% 
  filter(detail=="shut down") %>% 
  unique()

bars_reopen_baseline<- for_export %>% 
  filter(detail=="bars reopen") %>% 
  unique()
rest_reopen_baseline<- for_export %>% 
  filter(detail=="restaurants reopen") %>% 
  unique()

reclosed_baseline<- for_export %>% 
  filter(detail=="restaurants reclosed" | detail=="bars reclosed") %>% 
  unique()

reopened_again_baseline<- for_export %>% 
  filter(detail=="restaurants reopen again" | detail=="bars reopen again") %>% 
  unique()

# visualizing cases per capita and when bars/restaurants opened/closed
for_export %>% 
  select(-detail, -detail_date, -shutdown_baseline, -bars_reopen_baseline,-rest_reopen_baseline,
         -reclosed_baseline,-reopened_again_baseline) %>%
  unique() %>% 
  ggplot(aes(x=start_date, y=cases_pc)) +
  geom_col() +
  geom_vline(data=shutdown_baseline, aes(xintercept=start_date), alpha=.5, color="tomato") +
  geom_vline(data=bars_reopen_baseline, aes(xintercept=start_date), alpha=.5, color="midnightblue") +
  geom_vline(data=rest_reopen_baseline, aes(xintercept=start_date), alpha=.5, color="blue2") +
  geom_vline(data=reclosed_baseline, aes(xintercept=start_date), alpha=.5, color="tomato3") +
  geom_vline(data=reopened_again_baseline, aes(xintercept=start_date), alpha=.5, color="dodgerblue4") +
  facet_geo(~region) +
  labs(title="Covid-19 cases per capita and when bars/restaurants closed and reopened")

```
### After states shut down businesses

On the week of the shutdown versus 4 weeks later.

```{r after_shutdown, warning=F, message=F}
### shut down ----

shutdown <- for_export %>% 
  filter(shutdown_baseline %in% c(0, 3)) %>% 
  select(-detail, -detail_date) %>% 
  mutate(rest_per=rest_per*100, bar_per=bar_per*100) %>% 
  unique() %>% 
  select(region, NAME, MMWRyear, MMWRweek,cases, population) %>% 
  arrange(region, NAME, MMWRyear, MMWRweek) %>% 
  group_by(region, NAME, MMWRyear) %>% 
  mutate(id=row_number()) %>% 
  mutate(id=case_when(
    id == 1 ~ "shutdown",
    TRUE ~ "after"
  )) %>% 
  group_by(id) %>% 
  summarize(cases=sum(cases),
            population=sum(population)) %>% 
  mutate(cases_per_10k=cases/population*10000)


shutdown2 <- for_export %>% 
  filter(shutdown_baseline %in% c(0, 3)) %>% 
  select(-detail, -detail_date) %>% 
  mutate(rest_per=rest_per*100, bar_per=bar_per*100) %>% 
  unique() %>% 
  select(region, NAME, MMWRyear, MMWRweek,cases, population) %>% 
  arrange(region, NAME, MMWRyear, MMWRweek) %>% 
  group_by(region, NAME, MMWRyear) %>% 
  mutate(id=row_number()) %>% 
  mutate(id=case_when(
    id == 1 ~ "shutdown",
    TRUE ~ "after"
  )) %>% 
  mutate(per_capita=cases/population*10000) %>% 
  group_by(id) %>% 
  summarize(cases_per_10k=mean(per_capita))
  

shutdown_traffic <- for_export %>% 
  filter(shutdown_baseline %in% c(0, 3)) %>% 
  select(-detail, -detail_date) %>% 
  unique() %>% 
  select(region, NAME, MMWRyear, MMWRweek,rest_visits, bar_visits, total_devices) %>% 
  arrange(region, NAME, MMWRyear, MMWRweek) %>% 
  group_by(region, NAME, MMWRyear) %>% 
  mutate(id=row_number()) %>% 
  mutate(id=case_when(
    id == 1 ~ "shutdown",
    TRUE ~ "after"
  )) %>% 
  group_by(id) %>% 
  summarize(rest_visits=sum(rest_visits),
            bar_visits=sum(bar_visits),
            total_devices=mean(total_devices)) %>% 
  mutate(rest_traffic=rest_visits/total_devices*1000,
         bar_traffic=bar_visits/total_devices*1000)

shutdown <- left_join(shutdown, shutdown_traffic) %>% 
  select(id, cases, cases_per_10k, rest_traffic, bar_traffic) %>% 
  arrange(desc(id)) 

kable(shutdown)

shutdown_table_cases <- for_export %>% 
  filter(shutdown_baseline %in% c(0,4)) %>% 
  select(-detail, -detail_date) %>% 
  mutate(rest_per=rest_per*100,
         bar_per=bar_per*100) %>% 
  unique() %>% 
  select(region, NAME, MMWRyear, MMWRweek,cases_pc) %>% 
  arrange(region, NAME, MMWRyear, MMWRweek) %>% 
  group_by(region, NAME, MMWRyear) %>% 
  mutate(id=row_number()) %>% 
  mutate(id=case_when(
    id == 1 ~ "shutdown",
    TRUE ~ "after"
  )) %>% 
  ungroup() %>% 
  select(-MMWRweek) %>% 
  pivot_wider(names_from="id",values_from="cases_pc") %>% 
  mutate(more=case_when(
    after>shutdown ~ "More", 
    TRUE ~ "Less"))
```

How many states saw an increase or decrease in cases per capita 4 weeks later?

```{r restaurants_percap}
shutdown_table_cases %>% count(more) %>% 
kable()



```


### After states reopened bars

On the week of the reopening versus 4 weeks later.

```{r before_after_bars, warning=F, message=F}

nat_bars <- for_export %>% 
  filter(bars_reopen_baseline %in% c(0, 3)) %>% 
  select(-detail, -detail_date) %>% 
  mutate(bar_per=bar_per*100) %>% 
  unique() %>% 
  select(region, NAME, MMWRyear, MMWRweek, cases, population) %>% 
  arrange(region, NAME, MMWRyear, MMWRweek) %>% 
  group_by(region, NAME, MMWRyear) %>% 
  mutate(id=row_number()) %>% 
  mutate(id=case_when(
    id == 1 ~ "open",
    TRUE ~ "after"
  )) %>% 
  group_by(id) %>% 
  summarize(cases=sum(cases),
            population=sum(population)) %>% 
  mutate(cases_per_10k=cases/population*10000)

nat_bar_traffic <- for_export %>% 
  filter(bars_reopen_baseline %in% c(0, 3)) %>% 
  select(-detail, -detail_date) %>% 
  unique() %>% 
  select(region, NAME, MMWRyear, MMWRweek,bar_visits, total_devices) %>% 
  arrange(region, NAME, MMWRyear, MMWRweek) %>% 
  group_by(region, NAME, MMWRyear) %>% 
  mutate(id=row_number()) %>% 
  mutate(id=case_when(
    id == 1 ~ "open",
    TRUE ~ "after"
  )) %>% 
  group_by(id) %>% 
  summarize(bar_visits=sum(bar_visits),
            total_devices=mean(total_devices)) %>% 
  mutate(bar_traffic=bar_visits/total_devices*1000)

nat_bars <- left_join(nat_bars, nat_bar_traffic) %>% 
  select(id, cases, cases_per_10k, bar_traffic) %>% 
  arrange(desc(id))

kable(nat_bars)
```

How many states saw an increase or decrease in cases per capita 4 weeks later?

```{r before_after_bars2, warning=F, message=F}
bars_table_cases <- for_export %>% 
  filter(bars_reopen_baseline %in% c(0, 3)) %>% 
  select(-detail, -detail_date) %>% 
  mutate(bar_per=bar_per*100) %>% 
  unique() %>% 
  select(region, NAME, MMWRyear,MMWRweek, cases_pc) %>% 
  arrange(region, NAME, MMWRyear, MMWRweek) %>% 
  group_by(region, NAME, MMWRyear) %>% 
  mutate(id=row_number()) %>% 
  mutate(id=case_when(
    id == 1 ~ "open",
    TRUE ~ "after"
  )) %>% 
  ungroup() %>% 
  select(-MMWRweek) %>% 
  pivot_wider(names_from="id",values_from="cases_pc") %>% 
  mutate(more=case_when(
    after>open ~ "More", 
    TRUE ~ "Less"))

bars_table_cases %>% count(more) %>% 
kable()

```


### After states reopened restaurants

On the week of the reopening versus 4 weeks later.


```{r restaurants_reopened, warning=F, message=F}
#### national restaurants ---

nat_rest <- for_export %>% 
  filter(rest_reopen_baseline %in% c(0, 3)) %>% 
  select(-detail, -detail_date) %>% 
  mutate(rest_per=rest_per*100) %>% 
  unique() %>% 
  select(region, NAME, MMWRyear, MMWRweek,cases, population) %>% 
  arrange(region, NAME, MMWRyear, MMWRweek) %>% 
  group_by(region, NAME, MMWRyear) %>% 
  mutate(id=row_number()) %>% 
  mutate(id=case_when(
    id == 1 ~ "open",
    TRUE ~ "after"
  )) %>% 
  group_by(id) %>% 
  summarize(cases=sum(cases),
            population=sum(population)) %>% 
  mutate(cases_per_10k=cases/population*10000)

nat_rest_traffic <- for_export %>% 
  filter(rest_reopen_baseline %in% c(0, 3)) %>% 
  select(-detail, -detail_date) %>% 
  unique() %>% 
  select(region, NAME, MMWRyear, MMWRweek,rest_visits, total_devices) %>% 
  arrange(region, NAME, MMWRyear, MMWRweek) %>% 
  group_by(region, NAME, MMWRyear) %>% 
  mutate(id=row_number()) %>% 
  mutate(id=case_when(
    id == 1 ~ "open",
    TRUE ~ "after"
  )) %>% 
  group_by(id) %>% 
  summarize(rest_visits=sum(rest_visits),
            total_devices=mean(total_devices)) %>% 
  mutate(rest_traffic=rest_visits/total_devices*1000)

nat_rest <- left_join(nat_rest, nat_rest_traffic) %>% 
  select(id, cases, cases_per_10k, rest_traffic) %>% 
  arrange(desc(id))

kable(nat_rest)
```

How many states saw an increase or decrease in cases per capita 4 weeks later?

```{r restaurants_reopened2}
rest_table_cases <- for_export %>% 
  filter(rest_reopen_baseline %in% c(0, 3)) %>% 
  select(-detail, -detail_date) %>% 
  mutate(rest_per=rest_per*100) %>% 
  unique() %>% 
  select(region, NAME, MMWRyear, MMWRweek,cases_pc) %>% 
  arrange(region, NAME, MMWRyear, MMWRweek) %>% 
  group_by(region, NAME, MMWRyear) %>% 
  mutate(id=row_number()) %>% 
  mutate(id=case_when(
    id == 1 ~ "open",
    TRUE ~ "after"
  )) %>% 
  ungroup() %>% 
  select(-MMWRweek) %>% 
  pivot_wider(names_from="id",values_from="cases_pc") %>% 
  mutate(more=case_when(
    after>open ~ "More", 
    TRUE ~ "Less"))


rest_table_cases %>% count(more) %>% 
kable()

```




### After states reclosed restaurants and bars

On the week of the reopening versus 4 weeks later.


```{r restaurants_reclosed, warning=F, message=F}
#### national restaurants ---

nat_reclosed <- for_export %>% 
  filter(reclosed_baseline  %in% c(0, 3)) %>% 
  select(-detail, -detail_date) %>% 
  mutate(rest_per=rest_per*100, bar_per=bar_per*100) %>% 
  unique() %>% 
  select(region, NAME, MMWRyear, MMWRweek,cases, population) %>% 
  arrange(region, NAME, MMWRyear, MMWRweek) %>% 
  group_by(region, NAME, MMWRyear) %>% 
  mutate(id=row_number()) %>% 
  mutate(id=case_when(
    id == 1 ~ "reclosed",
    TRUE ~ "after"
  )) %>% 
  group_by(id) %>% 
  summarize(cases=sum(cases),
            population=sum(population)) %>% 
  mutate(cases_per_10k=cases/population*10000)

nat_reclosed_traffic <- for_export %>% 
  filter(reclosed_baseline  %in% c(0, 3)) %>% 
  select(-detail, -detail_date) %>% 
  unique() %>% 
  select(region, NAME, MMWRyear, MMWRweek,rest_visits, bar_visits, total_devices) %>% 
  arrange(region, NAME, MMWRyear, MMWRweek) %>% 
  group_by(region, NAME, MMWRyear) %>% 
  mutate(id=row_number()) %>% 
  mutate(id=case_when(
    id == 1 ~ "reclosed",
    TRUE ~ "after"
  )) %>% 
  group_by(id) %>% 
  summarize(rest_visits=sum(rest_visits),
            bar_visits=sum(bar_visits),
            total_devices=mean(total_devices)) %>% 
  mutate(rest_traffic=rest_visits/total_devices*1000,
         bar_traffic=bar_visits/total_devices*1000)

nat_reclosed <- left_join(nat_reclosed, nat_reclosed_traffic) %>% 
  select(id, cases, cases_per_10k, rest_traffic, bar_traffic) %>% 
  arrange(desc(id))

kable(nat_reclosed)
```

How many states saw an increase or decrease in cases per capita 4 weeks later?

```{r restaurants_reclosed2}
reclosed_table_cases <- for_export %>% 
  filter(reclosed_baseline %in% c(0, 3)) %>% 
  select(-detail, -detail_date) %>% 
  mutate(rest_per=rest_per*100) %>% 
  unique() %>% 
  select(region, NAME, MMWRyear, MMWRweek,cases_pc) %>% 
  arrange(region, NAME, MMWRyear, MMWRweek) %>% 
  group_by(region, NAME, MMWRyear) %>% 
  mutate(id=row_number()) %>% 
  mutate(id=case_when(
    id == 1 ~ "reclosed",
    TRUE ~ "after"
  )) %>% 
  ungroup() %>% 
  select(-MMWRweek) %>% 
  pivot_wider(names_from="id",values_from="cases_pc") %>% 
  mutate(more=case_when(
    after>reclosed ~ "More", 
    TRUE ~ "Less"))


reclosed_table_cases %>% count(more) %>% 
kable()

```